Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX Format\titlenote{(Produces the permission block, and copyright information). For use with SIG-ALTERNATE.CLS. Supported by ACM.}

### 2015-07-31 sig-alternate.tex
Set-up and Approach drafts

\subsection{Approach}
Defining accurately the human field of view in the implementation is a must. Holmqvist in \cite{holmqvist2011eye} specifies that the visual human range is $ \pm  40\degree $ in the horizontal and $ \pm 25\degree $ in the vertical. In \cite{walker1980clinical} Spector provides a more detailed specification for each eye splitting the vertical range into $ 60\degree $ the upper region and $ 75\degree $ the lower one. Moreover, the horizontal range gets separated in $ 60\degree $ inwards (towards the nose) and $ 95\degree $ outwards. In the present implementation, the first approach explained has been chosen representing the field of view using a cone with such dimensions.

We used a commercial RGB camera acquiring images of 640x480 pixels at 25 fps. The camera was located at 5 cm from the center of Nao's feet and its holder was tied to a wood base for stability purposes. Thus, the camera's objective was 9 cm above and $ 40\degree $ inclination with respect to the surface of the table. 
The subjects were typically located 50 cm away from the robot with the first tablet in front and the second one 30 cm to the left of the first one. In the same way, the experimenter was located around 60 cm to the left of the subject. Finally, two observers were located far away from the interaction field to manually assess

in the context of the CoWriter project \footnote{The primary repository is
\url{https://github.com/chili-epfl/cowriter_letter_learning}.}

\subsection{Experimental set-up}
Figure~\ref{drawSetup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{drawSetup}
    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
        nao} robot.  The robot writes on the tactile tablet, the child then
        corrects the robot by directly overwriting its letters on the tablet
        with a stylus. An adult (either a therapist or an experimenter,
        depending on the studies), remains next to the child to guide the work
        (prompting, turn taking, etc.). A second tablet allows to choose words
        and the camera captures the subject's head. In red, the focuses of attention
        in the scene.}
    \label{drawSetup}
\end{figure}

@@

As we observe the current set-up allow us to intuitively identify several focuses of attention along the interaction such as the two tablets, the robot, the experimenter and the observers located in the diagonal.

\subsection{From face detection to Focus of Attention}

@@

%equations

@@

1. How much the head pose contributes to the real gaze of the subject?
2. Which accuracy can we achieve during a real case scenario in terms of head pose estimation tracking?  
3. How accurate can we predict where the subject is looking at based only on the head pose estimation?
4. How can we evaluate the performance of the children based on the activity?

@@

\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Typical child-robot interaction scenario during the experiments}

    \label{realSetup}
\end{figure}

@@

\section{Discussion}
The results has shown that the percentage of accuracy in the tracking is not only conditioned by the subject's amount of movement, but also by the speed of the same movements: Restless children externalize greater levels of movement decreasing the tracker accuracy. Additionally, several classic computer vision challenges such as the lighting conditions, occlusions and the face angle with respect to the camera are factors that contribute to reduce the accuracy.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

@@

\subsection{Figures}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

\begin{figure*}
\centering

\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

@@

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

>>>>

The present studies has been developed in the context of the CoWriter project \footnote{The primary repository is \url{https://github.com/chili-epfl/cowriter_letter_learning}.}.

\subsection{Approach}

The first step is to estimate the head pose using the pin-hole model shown in equation \ref{eq:pinHole}. As we mention below dlib library facilitates the task by providing the 2D points \textit{m'} that composes the face silhouette. However, it is enough with obtaining eight spread points, such as right and left eyes, right and left tragions, sellion or nasal depression zone, pronasale or center of the nose, stomion or center of the mouth and menton.

Likewise, the 3D face model is needed. The proposed head pose estimation method would work best with a 3D face model of the query subject itself. When a 3D face model of the subject is not available, as in many practical situations, a 3D face model of a subject can be used. This is an estimation of the same anthropometric points \textit{M'} (for North American Caucasians) in the 3D space \cite{farkas1994anthropometry}. Moreover, the intrinsic camera parameters are necessary to be able to compose the camera matrix \textit{A}. The components are expanded in equation \ref{eq:pinHole2}.

\begin{equation}
s  \; m' = A [R|t] M'
\label{eq:pinHole}
\end{equation}
\begin{equation}
s
\begin{bmatrix}
x \\
y \\
1
\end{bmatrix}
=
\begin{bmatrix}
f_x & 0 & c_x  \\
0 & f_y & c_y  \\
0 & 0 & 1 
\end{bmatrix}
\begin{bmatrix}
r_{11} & r_{12} & r_{13} & t_x  \\
r_{21} & r_{22} & r_{23} & t_y  \\
r_{31} & r_{32} & r_{33} & t_z  
\end{bmatrix}
\begin{bmatrix}
X \\
Y \\
Z \\
1
\end{bmatrix}
\label{eq:pinHole2}
\end{equation}

Finally, the matrix of extrinsic parameters [R|t] is obtained by joining the rotation matrix R (calculated from the rotation vector) and translation vector \textit{t}. The result is a matrix projection \textbf{P} which maps a point in the 3D space onto a point in the 2D image place solving the equation \ref{eq:p}.

\begin{equation}
s  \; m' = P \; M'
\label{eq:p}
\end{equation}

The second step is to use the matrix [R|t] to set the correct translation and orientation of the \textit{tf} frame which represents the face in the 3D space. In order to achieve that it is necessary first, to set the origin defined by the vector \textit{t}. Second, the matrix \textit{R} needs to be reformulated into a quaternion as shown in equation \ref{eq:quat} due to the smoothly interpolation between them.

\begin{equation}
\begin{bmatrix}
qw \\
qx \\
qy \\
qz
\end{bmatrix}
=
\begin{bmatrix}
\sqrt{1 + r_{11} + r_{22} + r_{33}} /2 \\
(r_{32} - r_{23})/( 4 \cdot qw) \\
(r_{13} - r_{31})/( 4 \cdot qw) \\
(r_{21} - r_{12})/( 4 \cdot qw)
\end{bmatrix}
\label{eq:quat}
\end{equation}
\\
However, it is necessary to consider the cases where the division is performed by 0 (in 180\degree rotation about the y-axis for instance) or when \textit{qw} becomes close to zero.

The third step is to generate a representation of the field of view which origin and orientation are the ones defined by the face detection \textit{tf} frame.

Defining accurately the human field of view in the implementation is a must. Holmqvist in \cite{holmqvist2011eye} specifies that the visual human range is $ \pm  40\degree $ in the horizontal and $ \pm 25\degree $ in the vertical. In \cite{walker1980clinical} Spector provides a more detailed specification for each eye splitting the vertical range into $ 60\degree $ the upper region and $ 75\degree $ the lower one. Moreover, the horizontal range gets separated in $ 60\degree $ inwards (towards the nose) and $ 95\degree $ outwards. In the present implementation, the first approach explained has been chosen representing the field of view using a cone with such dimensions.

Assuming that the \textit{tf} frames representing the robotic agent, the two tablets, the experimenter and the observers are static, need to be monitored. This task consists on evaluating the intersection between the \textit{tf} frames and the field of view: If one focus of attention is within the region defined by the field of view, we assume the subject is looking at it. In order to solve such operation mathematically it is necessary to compute the transformation matrix between the coordinate frames to be able to locate the monitored frames \textit{A} in the subject's coordinate system \textit{B} (see equation \ref{eq:transform}).

\begin{equation}
v' = B(A^-1)v
\end{equation}

where \textit{v} is a point in \textit{A} that becomes v' in coordinate system \textit{B}. However, in practice there are, at least, three different \textit{tf} frames involved in the transformation (subject-camera, camera-base, base-focus). Following the transformation result, the distance of the monitored frame with respect to the x-axis face \textit{tf} frame, $ d_{x_{axis}} $, representing the subject's face can be computed as simple as in equation \ref{eq:pitagoras}. Always considering the x-axis as the one defining the face's orientation and thus field of view's main axis.

\begin{equation}
d_{x_{axis}} = \sqrt{t_y^2 + t_z^2}
\label{eq:pitagoras}
\end{equation}

It is necessary to compare the distance acquired $ d_{x_{axis}} $ with the radius of the field of view at the target \textit{tf} frame x-coordinate $ r_{fov} $ as shown in equation \ref{eq:fovtf}.

\begin{equation}
r_{fov} = tan\left(\frac{fov}{2}\right) \cdot t_x
\label{eq:fovtf}
\end{equation}

where \textit{fov} is the aperture angle of the field of view. Finally, if $ d_{x_{axis}}<r_{fov} $, the target is within the subject's field of view.

% Plot to clarify this??

\subsection{Experimental set-up}
Figure~\ref{realSetup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
            nao} robot.  The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting its letters on the tablet
            with a stylus. An adult (either a therapist or an experimenter,
            depending on the studies), remains next to the child to guide the work. 
            A second tablet allows to choose words and the camera captures the subject's face.}
    \label{realSetup}
\end{figure}

@@

@@

An RGB camera has been used to acquire images of 640x480 pixels at 25 fps. The camera was located at 5 cm from the center of Nao's feet and its holder was tied to a wood base for stability purposes. Thus, the camera's objective was 9 cm above and $ 40\degree $ inclination with respect to the surface of the table. 
The subjects were typically located 50 cm away from the robot with the first tablet in front and the second one 30 cm to the left of the first one. In the same way, the experimenter was located around 60 cm to the left of the subject. Finally, two observers were located far away from the interaction field to manually assess the state of the interaction.

Figure \ref{drawSetup} allows us to intuitively identify several focuses of attention along the interaction such as the two tablets, the robot, the experimenter and the observers located in the diagonal.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{drawSetup}
    \caption{\small Graphical representation of he set-up and the possible children focuses of attention in the scene, in red.}
    \label{drawSetup}
\end{figure}

\subsection{From face detection to Focus of Attention}
According to \cite{stiefelhagen2002tracking}

@@

@@

\begin{enumerate}
\item How much the head pose contributes to the real gaze of the subject?
\item Which percentage can we achieve during a real case scenario in terms of head pose estimation tracking?  
\item How accurate can we predict where the subject is looking at based only on the head pose estimation?
\item How can we evaluate the performance of the children based on the activity?
\end{enumerate}

@@

@@

\begin{table}
	\centering
	\caption{Tracking precision in face-to-face scenario}
	\begin{tabular}{|c|c|c|c|c|c|c|c|} \hline
	Subject & 1 & 2 & 3 & 4 & 5 & 6 & Avg\\ \hline
	Tracked & 78\% & \% & \% & \% & \% & \% & \% \\ \hline
	\end{tabular}
\end{table}

\section{Discussion}
The results has shown that the percentage of accuracy in the tracking is not only conditioned by the subject's amount of movement, but also by the speed of the same movements: Restless children externalize greater levels of movement decreasing the tracker accuracy. Additionally, several recurrent computer vision challenges such as the lighting conditions, occlusions and the face angle with respect to the camera are factors that contribute to reduce the accuracy.

\subsection{Tables}

@@

@@

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
